The Power of Two Random Choices

### 2014-01-07 article/article.tex
some additions to introdcution

\subsection{Balls into Bins}
\label{sec:ballsIntoBins}
Suppose that $m$ balls are sequentially thrown into $n$ bins, where each ball is placed in an independently and uniformly chosen random bin. We are interested in the \emph{number of balls in the fullest bin}, also called the \emph{maximum load}. Equivalently, one can state the \emph{gap} of the allocation, that is the difference of the number of balls in the fullest bin and the average amount of balls in the bins.

These \emph{balls into bins games} or \emph{allocation problems} have been studied extensively in the probability literature (see e.g., \cite{JK77}). When using the strategy as outlined above (that is, choose a bin independently and uniformly at random for each ball), the fullest bin contains $\Theta \left( \frac{\log n}{\log \log n} \right)$ for the case of $m = n$ with high probability \cite{RS98}. Since the average load of the bins is 1, this is also a bound for the gap. %More accurately, the number of balls in the bin with the most balls is $\Gamma^{-1}\left(n\right)\left(1+O\left(\frac{1}{\log \Gamma^{-1}\left(n\right)}\right)\right)$ \cite{G91}. %TODO: really include Gonnet? 
For the \emph{heavily loaded case}, where $m \gg n \log n$ the maximum load is $\frac{m}{n} + \Theta\left(\sqrt{\frac{m \log n}{n}}\right)$ with high probability \cite{RS98}. In the remainder of this paper, "heavily loaded case" will denote the case where $m = \omega\left(n\right)$ and the discussed strategy will be called \emph{once-choice balls into bins}.

Consider a variation of the above \emph{once-choice balls into bins} strategy, which we will call \emph{multiple-choice balls into bins}: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the \emph{least full bin}\footnote{If a tie occurs, chose one of the least full bins arbitrarily.}. Note that the $d$ selected bins are not necessarily distinct. In this variant, the maximum load is $\frac{m}{n} + \frac{\log \log n}{log d} \left( 1 + O\left(1\right)\right)$ for all $m$ with high probability. This was first proven by Azar et al. \cite{ABKU99} for the case $m = n$. (However, the case $d = 2$ was implicitly proven by Karp et al. \cite{KLM92}). See also \cite{RMS01} for a simplified proof and an overview of proof techniques for the given problem.  Unfortunately, the analysis breaks down for the heavily loaded case, where $m = \omega\left(n\right)$. Berenbrink at. all \cite{BFZR08} were the first to proof that the same bound holds for the heavily loaded case\footnote{However, they say that an event $\mathcal E$ occurs with high probability if $\Pr\left[\mathcal E \right]  \geq 1- n^{-\alpha}$, for an arbitrarily chosen constant $\alpha \geq 1$.}. A simplified proof may be found in \cite{TW13}.

>>>>

\subsection{Balls into Bins}
\label{sec:ballsIntoBins}
Suppose that $m$ balls are sequentially thrown into $n$ initially empty bins, where each ball is placed in an independently and uniformly chosen random bin. We are interested in the \emph{number of balls in the fullest bin}, also called the \emph{maximum load}. Equivalently, one can look at the \emph{gap} of the allocation, that is the difference of the number of balls in the fullest bin and the average amount of balls in the bins.

These \emph{balls into bins games} or \emph{allocation problems} have been studied extensively in the probability literature (see e.g., \cite{JK77}). Two basic variants of this problem exists (\cite{ABKU99}, \cite{RMS01}):
\begin{itemize}
\item In a \emph{static problem}, the balls are placed into the bins and no balls are ever removed.
\item In a \emph{dynamic problem}, balls can also be removed from the system. In a simple dynamic model, for each newly inserted ball a ball choosen uniformly and independently at random from among the balls currently in the bins is removed from the system.
\end{itemize}

When using the \emph{one-choice balls into bins} strategy as outlined above (that is, choose a bin independently and uniformly at random for each ball), the fullest bin contains $\Theta \left( \frac{\log n}{\log \log n} \right)$ for the case of $m = n$ w.h.p. \cite{RS98}. Since the average load of the bins is 1, this is also a bound for the gap. %More accurately, the number of balls in the bin with the most balls is $\Gamma^{-1}\left(n\right)\left(1+O\left(\frac{1}{\log \Gamma^{-1}\left(n\right)}\right)\right)$ \cite{G91}. %TODO: really include Gonnet? 
For the \emph{heavily loaded case}, where $m \gg n \log n$ the maximum load is $\frac{m}{n} + \Theta\left(\sqrt{\frac{m \log n}{n}}\right)$ with high probability \cite{RS98}. In the remainder of this paper, "heavily loaded case" will denote the case where $m = \omega\left(n\right)$.

Consider a variation of the above \emph{once-choice balls into bins} strategy, which we will call \emph{multiple-choice balls into bins}: For each ball, chose $d \geq 2$ bins independently and uniformly at random and place the ball into the \emph{least full bin}\footnote{If a tie occurs, chose one of the least full bins arbitrarily.}. Note that the $d$ selected bins are not necessarily distinct.

A \emph{uniform balls into bins problem} is a \emph{multiple choice balls and bins problem} where the $d$ random choices for one ball are independent and uniform. \emph{Non-uniform balls into bins problems} can reduce the maximum load and where first studied by Vöcking \cite{}.

In this variant, the maximum load is $\frac{m}{n} + \frac{\log \log n}{log d} \left( 1 + O\left(1\right)\right)$ for all $m$ with high probability. This was first proven by Azar et al. \cite{ABKU99} for the case $m = n$. (However, the case $d = 2$ was implicitly proven by Karp et al. \cite{KLM92}). See also \cite{RMS01} for a simplified proof and an overview of proof techniques for the given problem.  Unfortunately, the analysis breaks down for the heavily loaded case, where $m = \omega\left(n\right)$. Berenbrink at. all \cite{BFZR08} were the first to proof that the same bound holds for the heavily loaded case\footnote{However, they say that an event $\mathcal E$ occurs with high probability if $\Pr\left[\mathcal E \right]  \geq 1- n^{-\alpha}$, for an arbitrarily chosen constant $\alpha \geq 1$.}. A simplified proof may be found in \cite{TW13}.
